---
title: "Project EDA"
author: "Cathy C."
date-modified: "last-modified" 
execute:
  echo: true 
  eval: true 
  warning: false 
  freeze: true  
---

## 1 Load packages

```{r}
pacman::p_load(ggplot2, tidyverse, ggthemes, dplyr, tidyr, knitr)
```

## 2 Import file

```{r}
g2b_data <- read_csv("data/GovBIZ.csv")

kable(head(g2b_data))
```

## 3 Exploring Data

### 3.1 Check missing values

There is no missing values. Award date is character form, so **need to change to `date`.**

```{r}
#| code-fold: true
#| code-summary: "Show the code"
summary(g2b_data)

sum(!is.na(g2b_data$awarded_amt))
```

### 3.2 Number of agencies, suppliers

```{r}
#| code-fold: true
#| code-summary: "Show the code"
agency_count <- length(unique(g2b_data$agency))
supplier_count <- length(unique(g2b_data$supplier_name))

print(paste("# agency: ", agency_count))
print(paste("# supplier: ", supplier_count))

```

### 3.3 Awards amount

#### 3.3.1 Award amount in descending order (supplier)

```{r}
#| code-fold: true
#| code-summary: "Show the code"
library(dplyr)

desc_supp_award <- g2b_data %>%
  group_by(supplier_name) %>%
  summarise(total_award_amt = sum(awarded_amt, na.rm = TRUE)) %>%
  arrange(desc(total_award_amt)) %>%
  ungroup()
kable(head(desc_supp_award))

```

#### 3.3.2 Top 10 suppliers by total award amount

```{r}
#| code-fold: true
#| code-summary: "Show the code"
award_supp_top10 <- desc_supp_award %>%
  slice_head(n=10) %>%
  mutate(total_award_amt_mil = round(total_award_amt / 1000000, 4))
kable(award_supp_top10)
```

#### 3.3.3 Award amount in descending order (agency)

```{r}
#| code-fold: true
#| code-summary: "Show the code"
desc_agency_award <- g2b_data %>%
  group_by(agency) %>%
  summarise(total_award_amt = sum(awarded_amt, na.rm = TRUE)) %>%
  arrange(desc(total_award_amt)) %>%
  ungroup()
kable(head(desc_agency_award))
```

#### 3.3.4 Top 10 agencies by total award amount

```{r}
#| code-fold: true
#| code-summary: "Show the code"
award_agency_top10 <- desc_agency_award %>%
  slice_head(n=10) %>%
  mutate(total_award_amt_mil = round(total_award_amt / 1000000, 4))
kable(award_agency_top10)
```

#### 3.3.5 Top 10 suppliers from top 10 agencies by award amount

```{r}
#| code-fold: true
#| code-summary: "Show the code"
top_agencies_award <- g2b_data %>%
  group_by(agency) %>%
  summarise(total_awarded_amt = sum(awarded_amt, na.rm = TRUE), .groups = "drop") %>%
  slice_max(order_by = total_awarded_amt, n = 10)

#filter only the top 10 agencies and get top 10 suppliers for each
top_suppliers_award <- g2b_data %>%
  filter(agency %in% top_agencies_award$agency) %>%
  
  group_by(agency, supplier_name) %>%
  summarise(total_award_amt = sum(awarded_amt, na.rm = TRUE), .groups = "drop") %>%
  
  group_by(agency) %>%
  slice_max(order_by = total_award_amt, n = 10) %>%
  mutate(total_award_amt_mil = round(total_award_amt / 1000000, 4)) %>%
  ungroup()

kable(top_suppliers_award)
```

### 3.4 Tender count

#### 3.4.1 Tender count in descending order (supplier)

```{r}
#| code-fold: true
#| code-summary: "Show the code"
supp_tender_counts <- g2b_data %>%
  group_by(supplier_name) %>%
  summarise(tender_count = n(), .groups = "drop") %>%
  arrange(desc(tender_count))

kable(head(supp_tender_counts))
```

#### 3.4.2 Top 10 suppliers by total tender count

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tender_supp_top10 <- supp_tender_counts %>%
  slice_head(n=10) 

kable(tender_supp_top10)
```

#### 3.4.3 Tender count in descending order (agency)

```{r}
#| code-fold: true
#| code-summary: "Show the code"
agency_tender_counts <- g2b_data %>%
  group_by(agency) %>%
  summarise(tender_count = n(), .groups = "drop") %>%
  arrange(desc(tender_count))

kable(head(agency_tender_counts))
```

#### 3.4.4 Top 10 agencies by tender count

```{r}
#| code-fold: true
#| code-summary: "Show the code"
tender_agency_top10 <- agency_tender_counts %>%
  slice_head(n=10)

kable(tender_agency_top10)
```

#### 3.4.5 Top 10 supplies from top 10 agencies by tender count

```{r}
#| code-fold: true
#| code-summary: "Show the code"
top_agencies <- g2b_data %>%
  group_by(agency) %>%
  summarise(total_tenders = n(), .groups = "drop") %>%
  arrange(desc(total_tenders)) %>%
  slice_head(n = 10) %>%
  pull(agency)

# Get the top 10 suppliers for each of the top 10 agencies based on tender count
top_suppliers_per_agency <- g2b_data %>%
  filter(agency %in% top_agencies) %>%  
  group_by(agency, supplier_name) %>%
  summarise(tender_count = n(), .groups = "drop") %>%
  arrange(agency, desc(tender_count)) %>%  
  group_by(agency) %>%
  slice_head(n = 10) %>%  # Select top 10 suppliers per agency
  ungroup()

# View result
kable(top_suppliers_per_agency)
```

## 4 Unknown suppliers

```{r}
#| code-fold: true
#| code-summary: "Show the code"
unknown_data <- g2b_data %>%
  filter(supplier_name == "Unknown") 

unknown_count = nrow(unknown_data)
print(paste("Count of unknown suppliers:", unknown_count))

```

## 5 Top 10 agencies and their top 10 suppliers (by year)

```{r}
#| code-fold: true
#| code-summary: "Show the code"
library(dplyr)
library(lubridate)
g2b_data_backup <- g2b_data

# Convert award_date to Date and extract the year
g2b_data_backup$award_date <- dmy(g2b_data_backup$award_date)
g2b_data_backup$award_year <- year(g2b_data_backup$award_date)

# Check for missing dates and award_year values
missing_dates <- g2b_data_backup %>% filter(is.na(award_date))
missing_years <- g2b_data_backup %>% filter(is.na(award_year))
cat("Missing dates:", nrow(missing_dates), "\n")
cat("Missing award_years:", nrow(missing_years), "\n")

# Ensure there are no missing values for the year column
if (nrow(missing_years) > 0) {
  print("Rows with missing years:")
  print(missing_years)
} else {
  # Get top 10 agencies by total tender count per year
  top_agencies <- g2b_data_backup %>%
    group_by(award_year, agency) %>%
    summarise(total_tenders = n(), .groups = "drop") %>%
    arrange(award_year, desc(total_tenders)) %>%
    group_by(award_year) %>%
    slice_head(n = 10)  # Select top 10 agencies per year

  # Get top 10 suppliers for each of the top 10 agencies per year by tender count
  top_suppliers_per_agency <- g2b_data_backup %>%
    inner_join(top_agencies, by = c("award_year", "agency")) %>% # Filter to top agencies per year
    group_by(award_year, agency, supplier_name) %>%
    summarise(total_tenders = n(), .groups = "drop") %>%
    arrange(award_year, agency, desc(total_tenders)) %>%
    group_by(award_year, agency) %>%
    slice_head(n = 10)  # Select top 10 suppliers per agency per year

  # Split data into a list of tables, one for each year
  tables_by_year <- top_suppliers_per_agency %>%
    group_by(award_year) %>%
    group_split()
  
  unique_years <- unique(g2b_data_backup$award_year)
  unique_years <- sort(unique_years)

  # # Assign meaningful names to each table based on award_year
  if (length(tables_by_year) == length(unique_years)) {
    names(tables_by_year) <- as.character(unique_years)

    # Print each year's table
    for (year in names(tables_by_year)) {
      cat("\n\n### Top 10 Agencies & Suppliers for Year:", year, "\n")
      print(kable(tables_by_year[[year]]))
    }
  } else {
    print("Mismatch between unique years and the number of tables")
  }
}
```

::::: goals
::: goals-header
To do
:::

::: goals-container
1.  convert award_date data type
2.  extract year
3.  extract month
4.  remove unknown suppliers entry
:::
:::::
