---
title: "Hands-on_Ex_09"
author: "Cathy C."
date-modified: "last-modified" 
execute:
  echo: true 
  eval: true 
  warning: false 
  freeze: true  
---

# Modelling, Visualising and Analysing Network Data with R

## 1 Overview

Learning objectives include the following:

-   create graph object data frames, manipulate them using appropriate functions of *dplyr, lubridate, and tidygraph.*

-   build network graph visualisation using appropriate function of ***ggraph***

-   compute network geometrics using ***tidygraph***

-   build advanced gragh visualisation by incorporating the network ***geometrics***

-   build interactive network visualisation using ***visNetwork*** package

## 2 Getting started

Four network data modeling and visualisation packages will be installed and launched. They are **igraph, tidygraph, ggraph** and **visNetwork.** In addition, **tidyverse** and **lubridate** (specially designed to handle and wrangling time data) will be installed and launched, too.

## 3 The data

The data sets used in this hands-on exercise is from an oil exploration and extraction company. There are two data sets. One contains the nodes data and the other contains the edges (also know as link) data.

### 3.1 The edges data

-   *GAStech-email_edges.csv* which consists of two weeks of 9063 emails correspondances between 55 employees.

### 3.2 The nodes data

### 3.3 Importing network data from files

### 3.4 Reviewing the imported data

### 3.5 Wrangling time

### 3.6 Reviewomg the revised date fields

### 3.7 Wrangling attributes

### 3.8 Reviewing the revised edges file

### 

## 4 Creating network objects using **tidygraph**

## 5 Plotting static network graphs with **ggraph** package

## 6 Creating face graphs

## 7 Network metrics analysis

## 8 Building interactive network graph with visNetwork

## 9 Reference

# STUDY

```{r}
library(tidygraph)
create_ring(10)
```

```{r}
iris_clust <- hclust(dist(iris[1:4]))
iris_tree <- as_tbl_graph(iris_clust)
iris_tree
```

```{r}
iris_tree %>% activate(edges)
```

```{r}
as_tibble(iris_tree)
```

## The dplyr verbs

```{r}
iris_tree <- iris_tree %>% 
    activate(nodes) %>% 
    mutate(Species = ifelse(leaf, as.character(iris$Species)[label], NA)) %>% 
    activate(edges) %>% 
    mutate(to_setose = .N()$Species[to] == 'setosa')
iris_tree
```

`.N()` function is used to gain access to the node data while manipulating the edge data. Similarly `.E()` will give you the edge data and `.G()` will give you the `tbl_graph` object itself.

While one might expect all of `dplyr`s verbs to be supported in that manner, there is a clear limitation in the relational data structure that requires rows to maintain their identity. Thus, `summarise()` and `do()` are not allowed as there is no clear interpretation of how alterations on the node and edge data with these verbs should be interpreted. If these operations are required I suggest applying them to a tibble representation and then joining the result back in.

Speaking of joining, all joins from `dplyr` are supported. Nodes and edges are added and removed as required by the join. New edge data to be joined in must have a `to` and `from` column referencing valid nodes in the existing graph.

```{r}
library(dplyr)
iris_sum <- iris %>% 
    group_by(Species) %>% 
    summarise_all(mean) %>% 
    ungroup()
iris_tree <- iris_tree %>% 
    activate(nodes) %>% 
    left_join(iris_sum)
iris_tree
```

```{r}
library(ggraph)
gr1 <- create_notable('bull') %>% 
    mutate(name = letters[1:5])
gr2 <- create_ring(5) %>% 
    mutate(name = letters[4:8])

# Plot
gr1 %>% bind_graphs(gr2) %>% 
    ggraph(layout = 'kk') + 
    geom_edge_link() + 
    geom_node_point(size = 8, colour = 'steelblue') +
    geom_node_text(aes(label = name), colour = 'white', vjust = 0.4) + 
    ggtitle('Binding graphs') + 
    theme_graph()
```

```{r}
gr1 %>% graph_join(gr2) %>% 
    ggraph(layout = 'kk') + 
    geom_edge_link() + 
    geom_node_point(size = 8, colour = 'steelblue') +
    geom_node_text(aes(label = name), colour = 'white', vjust = 0.4) + 
    ggtitle('Joining graphs') + 
    theme_graph()
```
